### 2010-07-23 papers/icfp-2010-desc/paper_idesc.tex
Desc@ICFP: replace plusHost and leHost and +_H and \leq_H

An expression of this language is either a value, a conditional
expression, an addition of numbers, or a comparison of
numbers. Informally, their type is the following:
%
\[
\begin{array}{l@{\::\:\:}l}
\SYMBEcond           & \forall \Bhab{\V{ty}}{\Ty} . \Ebool \To \V{ty} \To \V{ty} \To \V{ty}  \\ 
\SYMBEplus           & \Enat \To \Enat \To \Enat                           \\
\SYMBEle             & \Enat \To \Enat \To \Ebool                          \\
\SYMBEval            & \forall \Bhab{\V{ty}}{\Ty} . \Val{\V{ty}} \To \V{ty}
\end{array}
\]
%
The function $\SYMBVal$, used in the definition of $\SYMBEval$, simply
maps a type $ty$ of the object language to the corresponding type in
the host language. Hence, the arguments of $\SYMBEval$ are ensured to
be of the expected type. We assume that $\Nat$ and $\Bool$ represent
natural numbers and booleans in the host language, equipped with an
addition operation $\F{plusHost}$ and a comparison function
$\F{leHost}$. We define $\SYMBVal$ as follows:
%
\[\stk{
\SYMBVal : \Ty \To \Set \\
\begin{array}{@{}l@{\:\mapsto\:\:}l}
\Val{\Enat}   & \Nat \\
\Val{\Ebool}  & \Bool
\end{array}
}\]

@@

Having implemented the syntax, we would like to describe its
semantics. To do so, we implement an evaluator. The type of the
evaluator is:
%
\[
\evalH : \PI{\V{ty}}{\Ty} 
         \IMu{\Ty}{\HExprD}{\V{ty}} \To
         \Val{\V{ty}}
\]
%
The type of $\evalH$ is strikingly similar to a
catamorphism. Indeed, implementing a single step of evaluation -- the
algebra -- is sufficient, as $\cataI$ gives, for free, the full
evaluator. The implementation is as follows:
%
\[\stk{
\evalOne : \PI{\V{ty}}{\Ty} \idescop{(\toIDesc{\HExprD})\:\V{ty}}{\Ty}{\SYMBVal}
           \To {\Val{\V{ty}}} \\
\begin{array}{@{}l@{}c@{}l@{\:\mapsto\:\:}l}
\evalOne\: & \_\: & \pair{\SYMBEval}{\V{x}}{}                                             & \V{x} \\
\evalOne\: & \_\: & \pair{\SYMBEcond}{\pair{\BoolTrue}{\pair{\V{x}}{\_}{}}{}}{}   & \V{x} \\
\evalOne\: & \_\: & \pair{\SYMBEcond}{\pair{\BoolFalse}{\pair{\_}{\V{y}}{}}{}}{}  & \V{y} \\
\evalOne\: & \Enat\: & \pair{\SYMBEplus}{\pair{\V{x}}{\V{y}}{}}{}                     & \F{plusHost}\: \V{x}\: \V{y} \\
\evalOne\: & \Ebool\: & \pair{\SYMBEle}{\pair{\V{x}}{\V{y}}{}}{}                      & \F{leHost}\: \V{x}\: \V{y} 
\end{array} \\
\\
\evalH : \PI{\V{ty}}{\Ty} 
           \IMu{\Ty}{\HExprD}{\V{ty}} \To
           \Val{\V{ty}} \\
\evalH\: \V{ty}\: \V{term} \mapsto \cataI_{\Ty} \:
                                 (\toIDesc{\HExprD})\: 
                                 \SYMBVal\: 
                                 \evalOne\: 
                                 \V{ty}\: 
                                 \V{term}
}\]

>>>>

\newcommand{\plusHost}{\mathop{\green{+_{\mathrm{H}}}}}
\newcommand{\leHost}{\mathop{\green{\leq_{\mathrm{H}}}}}

An expression of this language is either a value, a conditional
expression, an addition of numbers, or a comparison of
numbers. Informally, their type is the following:
%
\[
\begin{array}{l@{\::\:\:}l}
\SYMBEcond           & \forall \Bhab{\V{ty}}{\Ty} . \Ebool \To \V{ty} \To \V{ty} \To \V{ty}  \\ 
\SYMBEplus           & \Enat \To \Enat \To \Enat                           \\
\SYMBEle             & \Enat \To \Enat \To \Ebool                          \\
\SYMBEval            & \forall \Bhab{\V{ty}}{\Ty} . \Val{\V{ty}} \To \V{ty}
\end{array}
\]
%
The function $\SYMBVal$, used in the definition of $\SYMBEval$, simply
maps a type $ty$ of the object language to the corresponding type in
the host language. Hence, the arguments of $\SYMBEval$ are ensured to
be of the expected type. We assume that $\Nat$ and $\Bool$ represent
natural numbers and booleans in the host language, equipped with an
addition operation $\plusHost$ and a comparison function
$\leHost$. We define $\SYMBVal$ as follows:
%
\[\stk{
\SYMBVal : \Ty \To \Set \\
\begin{array}{@{}l@{\:\mapsto\:\:}l}
\Val{\Enat}   & \Nat \\
\Val{\Ebool}  & \Bool
\end{array}
}\]

@@

Having implemented the syntax, we would like to describe its
semantics. To do so, we implement an evaluator. The type of the
evaluator is:
%
\[
\evalH : \PI{\V{ty}}{\Ty} 
         \IMu{\Ty}{\HExprD}{\V{ty}} \To
         \Val{\V{ty}}
\]
%
The type of $\evalH$ is strikingly similar to a
catamorphism. Indeed, implementing a single step of evaluation -- the
algebra -- is sufficient, as $\cataI$ gives, for free, the full
evaluator. The implementation is as follows:
%
\[\stk{
\evalOne : \PI{\V{ty}}{\Ty} \idescop{(\toIDesc{\HExprD})\:\V{ty}}{\Ty}{\SYMBVal}
           \To {\Val{\V{ty}}} \\
\begin{array}{@{}l@{}c@{}l@{\:\mapsto\:\:}l}
\evalOne\: & \_\: & \pair{\SYMBEval}{\V{x}}{}                                             & \V{x} \\
\evalOne\: & \_\: & \pair{\SYMBEcond}{\pair{\BoolTrue}{\pair{\V{x}}{\_}{}}{}}{}   & \V{x} \\
\evalOne\: & \_\: & \pair{\SYMBEcond}{\pair{\BoolFalse}{\pair{\_}{\V{y}}{}}{}}{}  & \V{y} \\
\evalOne\: & \Enat\: & \pair{\SYMBEplus}{\pair{\V{x}}{\V{y}}{}}{}                     & \V{x} \plusHost \V{y} \\
\evalOne\: & \Ebool\: & \pair{\SYMBEle}{\pair{\V{x}}{\V{y}}{}}{}                      & \V{x} \leHost \V{y} 
\end{array} \\
\\
\evalH : \PI{\V{ty}}{\Ty} 
           \IMu{\Ty}{\HExprD}{\V{ty}} \To
           \Val{\V{ty}} \\
\evalH\: \V{ty}\: \V{term} \mapsto \cataI_{\Ty} \:
                                 (\toIDesc{\HExprD})\: 
                                 \SYMBVal\: 
                                 \evalOne\: 
                                 \V{ty}\: 
                                 \V{term}
}\]
